Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Comparing Hotspot Evolution in Two and Three Dimensions

### 2023-02-20 xrb3/paper.tex
new figures in place
add some more figs
some fig adjustments
some more details on resolution

\documentclass[preprint,times,tighten]{aastex631}

@@

In \citet{harpole:2021}, we explored several different rotation rates
and initial model thermal structures.  We found that a model with a
neutron star crust temperature of $3\times 10^8~\mathrm{K}$ resulted
in a flame that accelerated quickly as it spread across the neutron
star surface.  We adopt this model for the 3D simulation presented
here, since it requires the least amount of simulation time to yield a
developed flame.

Moving to 3D is very expensive, so to further reduce the computational
expense, we use a slightly lower resolution of 25 cm instead of 20 cm
from \citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.
Because the hotspot is placed in the center of the domain, the size of
the domain also becomes a constraint, and as a result, the evolution
time we can reach will become limited by the time it takes the burning
to approach the edge of the domain.  As a result, we focus here on the
early evolution.

Our 3D simulation used a base grid of $1024^2 \times 128$ zones with 2
levels of refinement, the first jumping by a factor of 4 and the next
by a factor of 2.  The domain has a size $(2.048\times
10^5~\mathrm{cm})^2 \times 2.56\times 10^4~\mathrm{cm}$, giving a fine
grid resolution of $25~\mathrm{cm}$.  The initial model and other
simulation parameters are identical to those used in
\citet{harpole:2021}.  We used static mesh refinement, to achieve
better load balancing, fully refining the atmosphere below a height of
$3600~\mathrm{cm}$.  These simulations used the 7-isotope He-burning
network used in \citet{eiden:2020}.  The simulations were run on the
OLCF Summit supercomputer, on 342 to 1366 nodes, with 6 NVIDIA V100
GPUs per node.  The entire computation was offloaded to GPUs following
the strategy described in \citet{castro_gpu}.  Overall, about 370,000
node-hours were used for the calculation.  The full state of the
calculation was output only every 0.005 s, as each output file is 2.7 TB
in size (for single precision data).  We output a few fields more
frequently for visualization.  We ran a 2D simulation with the same
resolution and refinement strategy to compare with here.

@@

Slice plot compared to 2-d simulation

Phase plot comparisons

>>>>

\documentclass[twocolumn,times,tighten]{aastex631}

@@

In \citet{harpole:2021}, we explored several different rotation rates
and initial model thermal structures.  We explored several different
temperature structures in our initial model.  We choose a model with a
neutron star crust temperature of $2\times 10^8~\mathrm{K}$.  We adopt
this model for the 3D simulation presented here, since it gives a clean
well-defined flame.

Moving to 3D is very expensive, so to further reduce the computational
expense, we use an aniostropic resolution, with 32 cm laterally and 16 cm resolution
vertically.  This is finer vertically than done in 
\citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.
Because the hotspot is placed in the center of the domain, the size of
the domain also becomes a constraint, and as a result, the evolution
time we can reach will become limited by the time it takes the burning
to approach the edge of the domain.  As a result, we focus here on the
early evolution.

Our 3D simulation used a base grid of $768^2 \times 192$ zones with 2
levels of refinement, the first jumping by a factor of 4 and the next
by a factor of 2.  The domain has a size $(1.96608\times
10^5~\mathrm{cm})^2 \times 2.4576\times 10^4~\mathrm{cm}$, giving a
fine grid resolution of $32~\mathrm{cm}$ laterally and
$16~\mathrm{cm}$ vertically.  The vertical resolution is finer to help
maintain hydrostatic equilibrium.  The initial model and other
simulation parameters are identical to those used in
\citet{harpole:2021}.  We used static mesh refinement, to achieve
better load balancing, fully refining the atmosphere below a height of
$3600~\mathrm{cm}$.  These simulations used the 7-isotope He-burning
network used in \citet{eiden:2020}.  The simulations were run on the
OLCF Summit supercomputer, on 342 to 1366 nodes, with 6 NVIDIA V100
GPUs per node.  The entire computation was offloaded to GPUs following
the strategy described in \citet{castro_gpu}.  Overall, about 250,000
node-hours were used for the calculation.  The full state of the
calculation was output only every 0.005 s, as each output file is 2.4 
TB in size (for single precision data).  We output a few fields more
frequently for visualization.  We ran a 2D simulation with the same
resolution and refinement strategy to compare with here.

@@

\begin{figure*}[t]
\centering
\plotone{time_series}
\caption{\label{fig:2d_abar} stuff}
\end{figure*}

\begin{figure}[t]
\centering
%\epsscale{0.75}
\plotone{abar_top_stack}
%\epsscale{1.0}
\caption{\label{fig:vr_abar} stuff}
\end{figure}

%% \begin{figure}[t]
%% \centering
%% \plotone{flame_wave_1000Hz_25cm_smallplt88081_Hnuc_annotated_top} \\
%% \plotone{flame_wave_1000Hz_25cm_smallplt174562_Hnuc_annotated_top.png}
%% \caption{\label{fig:vr_hnuc} stuff}
%% \end{figure}
